The symmetry of partner modelling

### 2016-02-04 mutual-modeling.tex
Updated section 'Studies'

@@

How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
If it was null, modelling would be interpreted as an individual process that
varies according to personality traits or cognitive skills: some persons tend to
model more accurately their peer while others don't care or can simply not
($\mathcal{H}_{1}$). If the correlation was close to 1, we could conclude that
the modelling accuracy is an emergent property of teams since there would be low
probability that, by chance, low accurate modellers are paired with accurate
modellers and conversely ($\mathcal{H}_{2}$). Our correlation is between these
two extremes but the very notion of significant intra-class correlation tend to
indicate that one may not consider peers as independent subjects, which is an
argument for the \emph{interaction} hypothesis $\mathcal{H}_{3}$.

@@

\subsubsection*{Subjects}
%ANON
%Ninety EPFL students participated in this experiment. We only selected EPFL
%students since knowledge of the campus geography had an impact both on group
Ninety students participated in this experiment. We only selected students from
our campus since knowledge of the campus geography had an impact both on group
performance and on mutual modelling: to represent the path of someone across
some space is difficult without an \textit{a priori} mental map of this space.
We formed groups of students who knew each other. We assigned 10 triads to each
of our three experimental conditions. Each condition was made up of
approximately 25\% of women, but we did not control gender repartition within
each triad.

@@

\paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
\gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more $A$
makes errors about $B$, the more $B$ does as well.

@@

This positive correlation observed in the symmetry question is surprising as
well: when acknowledging the high heterogeneity of spatial skills among
adults~\citep{liben1981spatial}, a significant positive correlation would
indicate that mutual modelling indeed stems from group processes more than
individual processes. The significant results on the triangle questions goes in
the same direction: either both the modeler and the modellee contribute to
accurate modelling, or the quality of teamwork has a global effect. Since team
members did not have many opportunities to directly interact during this
experiment, the factor that facilitates accurate mutual modelling is probably
more the overall group strategy than the quality of verbal interactions: a clear
strategy facilitates memorizing one's partner path.

@@

\paragraph{Symmetry question} We computed the absolute differences $\Delta_1$
between \gModel{A}{B} and \gModel{B}{A} over all pair of subjects within a triad
(3 values per triad, 20 triads), and compared them with the same differences
computed from random gradings (following the same grade distribution as for the
experimental data).

A t-test on the two sets releaved a significantly lower average difference in
the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
$t=-3.3, df=60, p = 0.0016$), supporting $\mathcal{H}_{2}$ or $\mathcal{H}_{3}$.

\paragraph{Triangle questions} $\Delta_2$ is computed in a similar way as the average of
the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
subjects. The average difference is 0.42 ($SD= 0.34$), and if not significantly
different from the same data obtained from random gradings ($t=-1.61, df=60,
p=0.11$). This rejects $\mathcal{H}_{2}$.

For $\Delta_3$, the average absolute difference between \gModel{C}{A} and
\gModel{C}{B} over the 60 subjects is 0.40 ($SD= 0.41$) and is significantly
lower than chance ($t=-2.62, df=60, p=0.01$): a given subject tends to model its
two partners with similar degrees of accuracy, which supports $\mathcal{H}_{1}$.
It could also support $\mathcal{H}_{3}$, but it does not since $\Delta_2$ is not
significant.

@@

\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
between self and social accuracy here because we do not have an estimation of
self-accuracy: subjects describe their emotions but we have no way to check if
there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
interpret this too far, but to show that there are large variations even within one
area  (perceiving emotions).

@@

The results evidence that richer media does not automatically foster better
modelling, on the contrary: groups using only the audio modality to exchange
their views were able to build more accurate models of each other than groups
using both video and audio, which is especially surprising when modelling
emotions: one would expect that video feedback indeed help building a better
model of the emotions of the other. Our interpretation is that viewing what
one's partner sees (shared graphical editor) was, in this task, more important
than seeing each other~\citep{gaver1993one,anderson1997impact}.

In this study, the accuracy of mutual modelling between two peers tend to be
symmetrical, which would support the idea that the quality of modelling is primarily
building upon the social interaction. This is however not fully supported by the
analysis of $\Delta_2$: $A$ being good at modelling $C$ does not correlate with
$B$ being good at modelling $C$.

On the other hand, $\Delta_3$ shows that a given $A$ tends to
model $B$ and $C$ at similar levels of accuracy: this evidences an individual
skill at modelling others (hypothesis $\mathcal{H}_{1}$).

@@

\paragraph{Rectangle question} The correlation between \gmodel{A}{A}
and \gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
significant in each condition: someone good at self-modelling is not necessarily
good at modelling someone else and vice-versa.

@@

In this experiment, the symmetry of mutual modelling is only found when subjects
are in conditions of symmetry of information. This consolidates the hypothesis
of \gModel{A}{B} being a property of interactions in context ($\mathcal{H}_{3}$)
rather than $A$'s cognitive attitude.

@@

The KAT has a significant effect on \gModel{A}{B}: peers more accurately
estimated their partners knowledge ($M = 1.11$) than those in the control
condition [$M = 0.98, t(1, 60) = 3.19, p < 0.01$, Cohen's $d = 0.83$, large
effect]. This is a trivial result since the KAT provided them with an initial
\gmodel{A}{B}. However, subjects have to predict the post-test score while the
KAT informed them about the pre-test score. Actually, pairs in the KAT condition
produced significantly more instances of 3 specific categories of interactions:
(1) elaborated utterances with rich contents,  (2) utterances asking about the
other's knowledge such as ``Did you understand how transmission works?'' (3)
utterances describing one's own knowledge (\gmodel{A}{A}) such as ``I don't
remember the Ranvier's thing...''. These 3 categories provide different account
of \gModel{A}{B}: (1) as an effect of the quality of interaction, (2) as an
effect of $A$'s effort to model $B$ or (3) as $B$'s effort to give cues to $A$ about his
own knowledge.

@@

\paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
0.05, ns$).

>>>>

We report five studies conducted  by different researchers in different contexts between 2000 and 2015. They do not form a consistent research strategy but actually the fact that some trends emerge despite their diversity constitutes indeed the richness of this line of work.

\begin{table*}[ht!]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
%\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
%$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
%$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
%$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

@@

How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}? It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. It is nonetheless far from 1 which implies we cannot discard the individual modelling skills. We collect more evidence along the next studies.

$\mathcal{H}_{3}$

%If it was null, modelling would be interpreted as an individual process that
%varies according to personality traits or cognitive skills: some persons tend to
%model more accurately their peer while others don't care or can simply not
%($\mathcal{H}_{1}$). If the correlation was close to 1, we could conclude that
%the modelling accuracy is an emergent property of teams since there would be low
%probability that, by chance, low accurate modellers are paired with accurate
%modellers and conversely ($\mathcal{H}_{2}$). Our correlation is between these
%two extremes but the very notion of significant intra-class correlation tend to
%indicate that one may not consider peers as independent subjects, which is an
%argument for the \emph{interaction} hypothesis $\mathcal{H}_{3}$.

@@

Ninety students participated in this experiment. We only selected students from
our campus since knowledge of the campus geography had an impact both on group
performance and on mutual modelling: to represent the path of someone across
some space is difficult without an \textit{a priori} mental map of this space.
We formed groups of students who knew each other. We assigned 10 triads to each
of our three experimental conditions. Each condition was made up of
approximately 25\% of women, but we did not control gender repartition within
each triad.

@@

\paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
\gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more $A$
makes errors about $B$, the more $B$ does as well (and vice-versa).

@@

This positive correlation observed in the symmetry question confirms the first study.  In this case, this is actually more surprising given  the high heterogeneity of spatial skills among adults~\citep{liben1981spatial}. This result supports  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The results regarding $\Delta_2$ may support both  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$  but discards   $\mathcal{H}_{1}$  (if the modeller's skill would matter a lot,  $\Delta_2$ would be lower). Conversely, $\Delta_3$  supports $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$ but discards   $\mathcal{H}_{2}$  (if the modeller's skill would matter a lot,  $\Delta_3$ would be lower). In summary, various indices support the 3 hypotheses, which implies there is some truth in each of them, but $\mathcal{H}_{3}$ is the only hypothesis that is not rejected by any index. We may hence, with great caution, concluded that the social perspective ($\mathcal{H}_{3}$) is moderately reinforced by this second study. Since the values of $\Delta_2$ and $\Delta_3$ are very similar, we do not interpret their minor difference as evidence for a stronger role of the modeller ($\mathcal{H}_{1}$) or the modellee ($\mathcal{H}_{1}$).

We have however to bring some nuances to the social viewpoint ($\mathcal{H}_{3}).  The main feature that can be associated to the team level in this experiment is probably not the quality of their verbal interactions per se (they interact mostly by drawing on a shared map), but rather the consistency of  the spatial exploration strategy: a clear strategy facilitates memorizing one's partner path.  One could argue whether the team strategy can be dissociated from the team interactions quality or constitues one of its components.

@@

\paragraph{Symmetry question} We computed the absolute differences $\Delta_1$
between \gModel{A}{B} and \gModel{B}{A} over all pair of subjects within a triad
(3 values per triad, 20 triads), and compared them with the same differences
computed from random gradings (following the same grade distribution as for the
experimental data). A t-test on the two sets revealed a significantly lower average difference in
the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
$t=-3.3, df=60, p = 0.0016$), which confirms the symmetry of mutual modelling.

\paragraph{Triangle questions} $\Delta_2$ is computed in a similar way as the average of
the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
subjects. The average difference is 0.42 ($SD= 0.34$), and is not significantly
different from the same index computed from random gradings ($t=-1.61, df=60,
p=0.11$).

For $\Delta_3$, the average absolute difference between \gModel{C}{A} and
\gModel{C}{B} over the 60 subjects is 0.40 ($SD= 0.41$) and is significantly
lower than chance ($t=-2.62, df=60, p=0.01$): a given subject tends to model its
two partners with similar degrees of accuracy.

@@

\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
between self and social accuracy here because we do not have an estimation of
self-accuracy: subjects describe their emotions but we have no way to check if
there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
interpret this too far, but to show that there are large variations even within one
area  (perceiving emotions). These variations shomehow question the existence of a general aptitude to model others ( $\mathcal{H}_{1}$).

@@

In this third study, the accuracy of mutual modelling  between two peers tend to be symmetrical, which  supports $\mathcal{H}_{2}$ and ($\mathcal{H}_{3}$.)  This result is contradicted by the fact that ($\Delta_2$) is not significant, which rejects both
 $\mathcal{H}_{2}$ and, to a lower extent,,  $\mathcal{H}_{3}$. Finally,  ($\Delta_3$) supports both  $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$.  In summary, this study brings some supports to  $\mathcal{H}_{3}$, but reveal again that $\mathcal{H}_{3}$ is only part of the explanation.

Like the previous one, this study leads us to refine what we mean by quality of interactions in $\mathcal{H}_{3}$. We  expected that the video channel would help peers building a more accurate model of the each other's emotions.  The results show the opposite: peers in the audio-only condition built more accurate models, probably because they concentrated more on the shared concept map. This confirms other studies that revealed that viewing what one's partner sees (shared graphical editor) is more important
than seeing each other~\citep{gaver1993one,anderson1997impact}. Therefore, what is meant by the quality of interactions is $\mathcal{H}_{3} is more than the linguistic features of dialogue but includes the way these interactions are articulated to the task.

@@

\paragraph{Rectangle question} The correlation between \gmodel{A}{A}
and \gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
significant in each condition: someone good at self-modelling is not necessarily
good at modelling someone else and vice-versa.  This seems to indicate that partner modelling
requires different skills than meta-cognition, despite their similarity at some level of abstraction.

@@

In this experiment, the symmetry of mutual modelling is  found but  in one condition, namely when subjects
 receive the same information before the task.  This condition corresponds to the situation tested in the three
previous studies and supports  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$.

How do we interpret the fact that the symmetry vanishes when peers receive different texts to read before the task? One explanation would be the difficult of mutual modelling when peers don't know what others read, but we found no significance of \gModel{A}{A} between conditions. Since texts were partly overlapping, another explanation is that, in absence of these initial common grounds, mutual modelling requires A to make evident to B what A thinks B does not know about A, which is the second level of modelling described in $\mathcal{H}_{2}$ . A low symmetry means that some peers are better than others at this second level of modelling, which supports the existence of such an individual skill, as stated in  $\mathcal{H}_{2}$

@@

The KAT has a significant effect on \gModel{A}{B}: peers more accurately
estimated their partners knowledge ($M = 1.11$) than those in the control
condition [$M = 0.98, t(1, 60) = 3.19, p < 0.01$, Cohen's $d = 0.83$, large
effect]. This is a trivial result since the KAT provided them with an initial
\gmodel{A}{B}. However, subjects have to predict the post-test score while the
KAT informed them about the pre-test score. Actually, pairs in the KAT condition
produced significantly more instances of 3 interesting categories of interactions:
 (1) utterances asking about the
other's knowledge such as ``Did you understand how transmission works?'' (2)
utterances describing one's own knowledge (\gmodel{A}{A}) such as ``I don't
remember the Ranvier's thing...'' and (3) elaborated utterances with rich contents,  These 3 categories provide different account
of \gModel{A}{B}: (1) as an
effect of $A$'s effort to model $B$  ($\mathcal{H}_{1}$), (2) as $B$'s effort to give cues to $A$ about his
own knowledge  ($\mathcal{H}_{2}$) and (3) as an effect of the quality of interaction  ($\mathcal{H}_{3}$) .

@@

\paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
0.05, ns$), which discards  ($\mathcal{H}_{2}$) and  ($\mathcal{H}_{3}$). In addition, the KAT supports the processing of modelling at the first level ($\mathcal{H}_{1}$) and not at the second level ($\mathcal{H}_{2}$). Hence the fact that the KAT enhances  \gModel{A}{B}  brings additional supports to $\mathcal{H}_{1}$. However, the analysis of verbal interactions find elements that actually supports the three hypotheses.

 ($\mathcal{H}_{3}$)
